Selection coefficients at the phylogenetic and population-genetic scale

### 2022-04-13 manuscript/main.tex
Supp. Mat.

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\avg}[1]{\left< #1 \right>} % for average
\newcommand{\Ne}{N_{\mathrm{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\mathrm{fix}}}
\newcommand{\Spop}{\beta}
\newcommand{\SpopMean}{\overline{\Spop}}
\newcommand{\Sphy}{S}
\newcommand{\SphyMean}{\overline{\Sphy}}

@@

    Adaptive evolution in proteins can be quantified by comparing sequence variation inside a population to the sequence divergence with a sister species\cite{mcdonald_adaptative_1991}.
    This method has been largely improved\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}, and it's application across many taxa has uncovered that a large proportion of substitutions are positively selected\cite{moutinho_variation_2019}.
    In practice, the estimated proportion of fixed substitutions that are positively selected varies from 30\% to 70\%.
    Inside a protein, structural properties such as solvent exposure are positively correlated with the proportion of positively selected substitutions\cite{moutinho_impact_2019}.
    The root causes for this widespread rate of positive selection are however scarcely investigated, often assumed to be the end result of change in the environment or in selection pressure driving individuals to adapt.
    In this view, positive selection is adaptive and the two terms are confounded, such that the underlying fitness landscape is changing with time, theoretically known as seascape~\cite{mustonen_fitness_2009}.
    However, even under the assumption that the fitness landscape is immutable, the nearly-neutral theory\cite{ohta_nearly_1992} predicts that positive selection can occur to compensate for deleterious substitutions that have accumulated along lineages.
    Theoretically, at the nearly-neutral equilibrium, the net flow of deleterious and advantageous mutations reaching fixation are equal~\cite{sella_application_2005}.
    In other words, mutations are more likely to be deleterious but are less likely to reach fixation, and the total number of fixation are compensated.
    In this view of positive selection, advantageous mutations are predictable and evolution is non-adaptive since individuals are not climbing toward higher fitness in the fitness landscape~\cite{lassig_predicting_2017}.
    Altogether, the goad is this study is to evaluate the extent of non-adaptive positive selection, by combining population-genetics and phylogenetics datasets, and ask a broad range of questions.
    Are proteins optimal?
    Is the fitness landscape stable across the mammalian evolution?
    Can we use phylogenetics to predict the selection coefficient of a mutation?
    Are deleterious mutations reliably purged only when selection overpowers drift?

@@

    % How to articulate phylogenetic and population genetic
    This information can be contrasted with the observed variation inside population, such that we can predict the scaled selection coefficient (the change in fitness) for each non-synonymous single nucleotide polymorphism (SNP) observed in a population (fig.~\ref{fig:method}).
    We developed and leveraged a pipeline integrating divergence and polymorphism data across the entire exome for 29 populations across 7 genera, namely \textit{Equus}, \textit{Canis}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo}.
    For each non-synonymous SNP observed at the population level, we predicted its scaled selection coefficient ($\Sphy$) in fitness from amino-acid fitness estimated at the mammalian scale (fig.~\ref{fig:homo-afr-results}A).
    Then we ask whether the variants that are supposedly advantageous are showing signature of short term evolution.
    In other words, if a variant is found in many mammals but not in ancestral humans, and a mutation in human is toward the mammal version, is this mutation advantageous for the individual carry it?
    All observed SNPs are divided in 5 classes of selection: severely deleterious with $\Sphy<-3$, deleterious with $-3<\Sphy<-1$, weakly deleterious with $-1<\Sphy<0$, weakly advantageous with $0<\Sphy<1<$ and advantageous with $1<\Sphy$.
    For each class of selection, we first derived its site-frequency spectrum (SFS), which represent the proportion of mutations with a given number of derived allele in the population (fig.~\ref{fig:homo-afr-results}B).
    Compared to synonymous mutations (supposedly neutral), mutations predicted to be advantageous are over-represented at higher frequency, while severally deleterious are under-represented at high frequency, a signature that our prediction are qualitatively in the right direction.
    Aside from the frequency of the derived allele, another information is the total number of observed alleles compared to its expectation, given by the total mutation rate.
    The total mutation rate for each class is computed from the ancestral genome reconstruction (fig.~\ref{fig:homo-afr-results}C).
    For example, for the category of selection with $1<\Sphy$, the expected total mutation rate is obtained by adding together mutation rates for all possible non-synonymous mutations for which the derived allele ($d$) is toward a fitter amino-acid than the ancestral allele ($a$) with $S > 1$.
    The total mutation rate and the SFS are combined to estimate the distribution of the selection coefficient given by population-based model~\cite{tataru_inference_2017, tataru_polydfe_2020}.
    These population based models cannot give a specific estimate of $\Spop$ for each SNP, but by gathering information across many SNPs, they can estimate the proportion of advantageous mutations $p_+=\proba \left( 1 < \Spop \right)$, of nearly-neutral mutations $p_0=\proba \left( -1 < \Spop < 1 \right)$ and of deleterious mutations $p_-=\proba \left( \Spop < -1 \right)$ (fig.~\ref{fig:homo-afr-results}D).

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure2}
        \caption{
            Panel A: Histogram of predicted selection coefficient (S) for each observed mutation in a sample of 512 individuals (1024 alleles) with African descent.
            Mutations are divided in 5 classes of selection: severely deleterious (blue), deleterious (green), weakly deleterious (light green), weakly advantageous (yellow) and advantageous (red).
            Panel B: Site-frequency spectrum (SFS) represent the proportion of mutations (y-axis) with a given number of derived allele in the population (x-axis).
            SFS are drawn for a random sample of 16 alleles (mean in solid line and standard deviation in filled color) for each class of selection and for synonymous mutations supposedly neutral (black).
            Supposedly advantageous are over-represented at higher frequency, while severally deleterious are under-represented at high frequency, a signature of selection.
            Panel C. Histogram of predicted selection coefficient (S) for each possible mutation away from the ancestral human genome, weighted by the mutation rate.
            For each class of selection, the sum of histogram in this class thus gives the expected total mutation rate toward this class.
            If they are less mutations observed than expected, this class is thus undergoing purifying selection.
            Panel D. For each class of selection (and for the set of all non-synonymous mutations), information from the SFS and the expected total mutation rate are combined at the population-genetic scale to estimate the proportion of advantageous mutations $p_+=\proba \left( 1 < \Spop \right)$, of nearly-neutral mutations $p_0=\proba \left( -1 < \Spop < 1 \right)$ and of deleterious mutations $p_-=\proba \left( \Spop < -1 \right)$.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure3}
        \caption{
            Reproducibility of results shown in figure~\ref{fig:homo-afr-results} in 29 populations across 7 genera.
            From a site-frequency spectrum and total mutation rates are combined at the population-genetic scale to estimate the proportion of advantageous mutations $p_+=\proba \left( 1 < \Spop \right)$ in top panel, of nearly-neutral mutations $p_0=\proba \left( -1 < \Spop < 1 \right)$ in middle panel and of deleterious mutations $p_-=\proba \left( \Spop < -1 \right)$ in bottom panel.
        }
        \label{fig:all-pop-results}
    \end{figure*}

    Because of this observed overlap between phylogenetic and population-genetic methods, both estimating a scaled selection coefficient, we could study the estimation of the mean $\SpopMean$ as a function of the mean $\SphyMean$.
    However, the mean selection coefficient estimated at the phylogenetic scale is usually drastically negative because of the heavy tail of the negative part of the estimated DFE.
    Hence we focused on the relationship between $\SpopMean$ and $\SphyMean$ for a subset of SNPs, for which $\Sphy$ is close to $0$.
    Sliding window of groups of $5.000$ SNPs are taken with a mean $\SphyMean$ between $-0.5$ and $0.5$ (fig.~\ref{fig:selcoeff-pop-phylo}A). 
    $\SpopMean$ is estimated for each group, with some groups that might have part of same SNPs that overlap.
    The relationship between $\Spop$ and $\SphyMean$ is best described as a second order polynomial, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\Spop/\Sphy$ for each population.
    We observ that $\Spop/\Sphy$, representing the efficiency of selection at the short versus long term is positively correlated with synonymous diversity, a proxy of effective population siz.
    Altogether, population with large effective population size are under more efficient selection.

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure4}
        \caption{
            Panel A. Mean selection coefficient at the population scale $\SpopMean$ as a function of phylogenetic $\SphyMean$.
            Observed SNPs are sorted and groups of $5.000$ SNPs are taken as a sliding windows with increasing mean $\SphyMean$.
            For each group, $\SpopMean$ (y-axis) is estimated based on SFS and the total mutation rate (between both extreme of $\SphyMean$ in the group) and compared to the mean $\Sphy$ (y-axis).
            A second order polynomial is fitted for each population, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ given in legend.
            Panel B. Ratio $\SpopMean/\SphyMean$ as a function of the synonymous diversity for each population.
            Populations with higher diversity are showing higher ratio of ratio $\SpopMean/\Sphy$.
        }
        \label{fig:selcoeff-pop-phylo}
    \end{figure*}d

@@

    \subsection{Selection coefficient ($\Spop$) in population-based method}
    \label{subsec:s-polymorphism-method}
    The probability to sample allele at a given frequency (before fixation or extinction) is informative of its scaled selection coefficient at the population-scale ($\Spop$).
    Pooled across many sites, the SFS is thus informative on the underlying $\Spop$ of mutations, given we have a neutral expectation.
    In this configuration, a single $\Spop$ for all sampled mutations is biologically not realistic.
    Accordingly, a distribution of fitness effects of mutations (DFE) is assumed, usually modelled as a continuous distribution~\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009}.
    In this study, we used the software polyDFE\cite{tataru_inference_2017, tataru_polydfe_2020}, for which the DFE ($\phi$) is given by a mixture of a $\Gamma$ and Exponential distributions, parameterized by $\Spop_d$ , $b$, $p_b$
    and $\Spop_b$ as:
    \begin{equation*}
        \phi \left( \Spop; \Spop_d , b, p_b, \Spop_b \right) =
        \begin{dcases}
            \left( 1 - p_b \right) f_{\Gamma}(-\Spop; -\Spop_d, b) & \text{ if $\Spop \leq 0$,} \\
            p_b f_{e}(\Spop; \Spop_b) & \text{ if $\Spop \leq 0$,} \\
        \end{dcases}
    \end{equation*}
    where $\Spop_d$ is the mean of the DFE for $\Spop \leq 0$
    $b$ is the shape of the $\Gamma$ distribution,
    $p_b$ is the probability that $\Spop > 0$,
    $\Spop_b$ is the mean of the DFE for $\Spop > 0$,
    and $f_{\Gamma}(x; m, b)$ is the density of the $\Gamma$ distribution with mean m and shape b, while $f_{e}(x; m)$ is the density of the Exponential distribution with mean $m$.
    Once the DFE is fitted to the data, the proportion of advantageous ($p_+$), nearly-neutral ($p_0$) and deleterious mutations ($p_-$) are given as:
    \begin{align*}
        p_+ &= \proba_{\phi} \left( 1 < \Spop \right) = p_b \int_{1}^{+\infty} f_{e}(\Spop; \Spop_b)  \\
        p_0 &= \proba_{\phi} \left( -1 < \Spop < 1 \right) = p_b \int_{0}^{1} f_{e}(\Spop; \Spop_b) + \left( 1 - p_b \right) \int_{0}^{1} f_{\Gamma}(\Spop; \Spop_d, b) \\
        p_- &= \proba_{\phi} \left( \Spop < -1 \right) = \left( 1 - p_b \right) \int_{1}^{+\infty} f_{\Gamma}(\Spop; \Spop_d, b)
    \end{align*}
    And finally, the mean of the DFE $\overline{\Spop}$ is given by:
    \begin{align*}
        \SpopMean & = \int_{-\infty}^{+\infty} \Spop \phi \left( \Spop; \Spop_d , b, p_b, \Spop_b \right) \\
                         & =  p_b \Spop_b - \left( 1 - p_b \right) \Spop_d
    \end{align*}
    PolyDFE requires one SFS for non-synonymous mutations and one for synonymous mutations (neutral expectation), as well as the total mutation rate (mutation rate per site multiplied by number of sites) on which each SFS has been sampled.
    The total mutation rate for each SFS is thus obtained using the following three-steps procedure.
    First, for each gene, the reference DNA sequence is adjusted with ancestral and fixed polymorphism for each population.
    Second, from this reconstructed sequence, all possible mutations are computed, weighted by the mutation rate between nucleotide estimated at the phylogenetic scale $\UniDimArray{\mu}$.
    Third, these possible mutation are then classified whether they are synonymous or non-synonymous (stop mutations are excluded), non-synonymous mutations are then also classified depending on their selection coefficient ($\Sphy$).
    From the SFS and the total mutation rate for the both synonymous and for the class of non-syonymous, polyDFE estimates the DFE $\phi$ using maximum likelihood.\@.

>>>>

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\avg}[1]{\left< #1 \right>} % for average
\newcommand{\Ne}{N_{\mathrm{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\mathrm{fix}}}
\newcommand{\Spop}{\beta}
\newcommand{\SpopMean}{\overline{\Spop}}
\newcommand{\Sphy}{S}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\polyDel}{\proba \left[ {\Spop < -1} \right]}
\newcommand{\polyNeutral}{\proba \left[ -1 < \Spop < 1 \right]}
\newcommand{\polyAdv}{\proba \left[ 1 < \Spop  \right]}

@@

    Adaptive evolution in proteins can be quantified by comparing sequence variation inside a population to the sequence divergence with a sister species\cite{mcdonald_adaptative_1991}.
    Indeed, strongly adaptive mutations are expected to be observed in substitutions but not in polymorphism because they reach fixation quickly.
    This method has been largely improved to account for possible bias-inducing factors such as the presence of weak positive selection, background selection and demography fluctuations\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    It's application across many taxa has uncovered that a large proportion of substitutions are positively selected\cite{moutinho_variation_2019}.
    In practice, the estimated proportion of fixed substitutions that are positively selected varies from 30\% to 70\% (reference).
    Inside a protein, structural properties such as solvent exposure are positively correlated with the proportion of positively selected substitutions\cite{moutinho_impact_2019}.
    The root causes for this widespread rate of positive selection are however scarcely investigated, often assumed to be the end result of change in the environment or in selection pressure driving individuals to adapt (reference).
    In this view, positive selection is adaptive, such that the underlying fitness landscape is changing with time in response to external constraints, theoretically known as seascape\cite{mustonen_fitness_2009}.
    However, even under the assumption that the fitness landscape is stable, the nearly-neutral theory\cite{ohta_nearly_1992} predicts that positive selection can occur to compensate for deleterious substitutions that have accumulated along lineages.
    Theoretically, at the nearly-neutral equilibrium, the net flow of deleterious and advantageous mutations reaching fixation are equal\cite{sella_application_2005}.
    In other words, mutations are more likely to be deleterious but are less likely to reach fixation, and the total number of fixation are compensated.
    Under this kind of positive selection, advantageous mutations are predictable and evolution is non-adaptive since there is no genetic innovation in response to external constraint, just an advantage to restore a good functioning of proteins\cite{lassig_predicting_2017}.

    Altogether, the goal of this study is to evaluate the extent of non-adaptive positive selection, by combining population-genetics and phylogenetics datasets, and ask a broad range of questions.
    Are proteins optimal?
    Is the fitness landscape stable across the mammalian evolution?
    Can we use phylogenetics to predict the selection coefficient of a mutation?
    Are deleterious mutations reliably purged only when selection overpowers drift?

@@

    % How to articulate phylogenetic and population genetic
    This information can be contrasted with the observed variation inside population, such that we can predict the scaled selection coefficient (the change in fitness) for each non-synonymous single nucleotide polymorphism (SNP) observed in a population (fig.~\ref{fig:method}).
    We developed and leveraged a pipeline integrating divergence and polymorphism data across the entire exome for 29 populations across 7 genera, namely \textit{Equus}, \textit{Canis}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo}.
    For each non-synonymous SNP observed at the population level, we predicted its scaled selection coefficient ($\Sphy$) in fitness from amino-acid fitness estimated at the mammalian scale (fig.~\ref{fig:homo-afr-results}A).
    Then we ask whether the variants that are supposedly advantageous are showing signature of short term evolution.
    In other words, if a variant is found in many mammals but not in ancestral humans, and a mutation in human is toward the mammal version, is this mutation advantageous for the individual carry it?
    All observed SNPs are divided in 5 classes of selection: severely deleterious with $\Sphy<-3$, deleterious with $-3<\Sphy<-1$, weakly deleterious with $-1<\Sphy<0$, weakly advantageous with $0<\Sphy<1<$ and advantageous with $1<\Sphy$.
    For each class of selection, we first derived its site-frequency spectrum (SFS), which represent the proportion of mutations with a given number of derived allele in the population (fig.~\ref{fig:homo-afr-results}B).
    Compared to synonymous mutations (supposedly neutral), mutations predicted to be advantageous are over-represented at higher frequency, while severally deleterious are under-represented at high frequency, a signature that our prediction are qualitatively in the right direction.
    Aside from the frequency of the derived allele, another information is the total number of observed alleles compared to its expectation, given by the total mutation rate.
    The total mutation rate for each class is computed from the ancestral genome reconstruction (fig.~\ref{fig:homo-afr-results}C).
    For example, for the category of selection with $1<\Sphy$, the expected total mutation rate is obtained by adding together mutation rates for all possible non-synonymous mutations for which the derived allele is toward a fitter amino-acid than the ancestral allele with $\Sphy > 1$.
    The total mutation rate and the SFS are combined to estimate the distribution of the selection coefficient $\Spop$ given by population-based model\cite{tataru_inference_2017, tataru_polydfe_2020}.
    These population based models cannot give a specific estimate of $\Spop$ for each SNP, but by gathering information across many SNPs, these models can estimate the proportion of advantageous mutations ($\polyAdv$), of nearly-neutral mutations ($\polyNeutral$) and of deleterious mutations ($\polyDel$) (fig.~\ref{fig:homo-afr-results}D).

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure2}
        \caption{
            Panel A: Histogram of predicted selection coefficient (S) for each observed mutation in a sample of 512 individuals (1024 alleles) with African descent.
            Mutations are divided in 5 classes of selection: severely deleterious (blue), deleterious (green), weakly deleterious (light green), weakly advantageous (yellow) and advantageous (red).
            Panel B: Site-frequency spectrum (SFS) represent the proportion of mutations (y-axis) with a given number of derived allele in the population (x-axis).
            SFS are drawn for a random sample of 16 alleles (mean in solid line and standard deviation in filled color) for each class of selection and for synonymous mutations supposedly neutral (black).
            Supposedly advantageous are over-represented at higher frequency, while severally deleterious are under-represented at high frequency, a signature of selection.
            Panel C. Histogram of predicted selection coefficient (S) for each possible mutation away from the ancestral human genome, weighted by the mutation rate.
            For each class of selection, the sum of histogram in this class thus gives the expected total mutation rate toward this class.
            If they are less mutations observed than expected, this class is thus undergoing purifying selection.
            Panel D. For each class of selection (and for the set of all non-synonymous mutations), information from the SFS and the expected total mutation rate are combined at the population-genetic scale to estimate the proportion of advantageous mutations $\polyAdv$, of nearly-neutral mutations $\polyNeutral$ and of deleterious mutations $\polyDel$.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure3}
        \caption{
            Reproducibility of results shown in figure~\ref{fig:homo-afr-results} in 29 populations across 7 genera.
            From a site-frequency spectrum and total mutation rates are combined at the population-genetic scale to estimate the proportion of advantageous mutations $\polyAdv$ in top panel, of nearly-neutral mutations $\polyNeutral$ in middle panel and of deleterious mutations $\polyDel$ in bottom panel.
        }
        \label{fig:all-pop-results}
    \end{figure*}

    Because of this observed overlap between phylogenetic and population-genetic methods, both estimating a scaled selection coefficient, we could study the estimation of the mean $\SpopMean$ as a function of the mean $\SphyMean$.
    However, the mean selection coefficient estimated at the phylogenetic scale is usually drastically negative because of the heavy tail of the negative part of the estimated DFE.
    Hence we focused on the relationship between $\SpopMean$ and $\SphyMean$ for a subset of SNPs, for which $\SphyMean$ is close to $0$.
    Sliding window of groups of $5.000$ SNPs are taken with a mean $\SphyMean$ between $-0.5$ and $0.5$ (fig.~\ref{fig:selcoeff-pop-phylo}A).
    $\SpopMean$ is estimated for each group, with some groups that might have part of same SNPs that overlap.
    The relationship between $\SpopMean$ and $\SphyMean$ is best described as a second order polynomial, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ for each population.
    We observ that $\SpopMean/\SphyMean$, representing the efficiency of selection at the short versus long term is positively correlated with synonymous diversity, a proxy of effective population siz.
    Altogether, population with large effective population size are under more efficient selection.

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure4}
        \caption{
            Panel A. For each population, proportion of advantageous ($\polyAdv$), nearly-neutral ($\polyNeutral$) and of deleterious ($\polyDel$) mutations estimated at the population-genetic scale.
            Estimation is performed with SNPs predicted to be weakly-advantageous at the phylogenetic scale $0<\Sphy<1<$.
            Panel B. Mean selection coefficient at the population scale $\SpopMean$ as a function of phylogenetic $\SphyMean$.
            Observed SNPs are sorted and groups of $5.000$ SNPs are taken as a sliding windows with increasing mean $-0.5 \leq \SphyMean \leq 0.5$.
            For each group along the sliding window, $\SpopMean$ (y-axis) is estimated based on SFS and the total mutation rate, and compared to the mean $\SphyMean$ (x-axis).
            A second order polynomial is fitted for each population, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ given in legend.
            Panel C. Ratio $\SpopMean/\SphyMean$ as a function of the synonymous diversity for each population.
            Populations with higher diversity are showing higher ratio of ratio $\SpopMean/\SphyMean$.
        }
        \label{fig:selcoeff-pop-phylo}
    \end{figure*}

@@

    \subsection{Selection coefficient ($\Spop$) in population-based method}
    \label{subsec:s-polymorphism-method}
    The probability to sample allele at a given frequency (before fixation or extinction) is informative of its scaled selection coefficient at the population-scale ($\Spop$).
    Pooled across many sites, the SFS is thus informative on the underlying $\Spop$ of mutations, given we have a neutral expectation.
    In this configuration, a single $\Spop$ for all sampled mutations is biologically not realistic.
    Accordingly, a distribution of fitness effects of mutations (DFE) is assumed, usually modelled as a continuous distribution\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009}.
    In this study, we used the software polyDFE\cite{tataru_inference_2017, tataru_polydfe_2020}, for which the DFE ($\phi$) is given by a mixture of a $\Gamma$ and Exponential distributions, parameterized by $\Spop_d$ , $b$, $p_b$
    and $\Spop_b$ as:
    \begin{equation*}
        \phi \left( \Spop; \Spop_d , b, p_b, \Spop_b \right) =
        \begin{dcases}
            \left( 1 - p_b \right) f_{\Gamma}(-\Spop; -\Spop_d, b) & \text{ if $\Spop \leq 0$,} \\
            p_b f_{e}(\Spop; \Spop_b) & \text{ if $\Spop > 0$,} \\
        \end{dcases}
    \end{equation*}
    where $\Spop_d < 0 $ is the mean of the DFE for $\Spop \leq 0$,
    $b > 0$ is the shape of the $\Gamma$ distribution,
    $0 \leq p_b \leq 1$ is the probability that $\Spop > 0$,
    $\Spop_b > 0$ is the mean of the DFE for $\Spop > 0$,
    and $f_{\Gamma}(x; m, b)$ is the density of the $\Gamma$ distribution with mean m and shape b, while $f_{e}(x; m)$ is the density of the Exponential distribution with mean $m$.
    Once the DFE is fitted to the data, the proportion of advantageous ($\polyAdv$), nearly-neutral ($\polyNeutral$) and deleterious mutations ($\polyDel$) are given as:
    \begin{align*}
        \polyAdv &= p_b \int_{1}^{+\infty} f_{e}(\Spop; \Spop_b) \der \Spop  \\
        \polyNeutral &= p_b \int_{0}^{1} f_{e}(\Spop; \Spop_b) \der \Spop + \left( 1 - p_b \right) \int_{0}^{1} f_{\Gamma}(\Spop; -\Spop_d, b) \der \Spop \\
        \polyDel &= \left( 1 - p_b \right) \int_{1}^{+\infty} f_{\Gamma}(\Spop; -\Spop_d, b) \der \Spop
    \end{align*}
    And finally, the mean of the DFE $\overline{\Spop}$ is given by:
    \begin{align*}
        \SpopMean & = \int_{-\infty}^{+\infty} \Spop \phi \left( \Spop; \Spop_d , b, p_b, \Spop_b \right) \der \Spop \\
        & =  p_b \Spop_b + \left( 1 - p_b \right) \Spop_d
    \end{align*}
    PolyDFE requires one SFS for non-synonymous mutations and one for synonymous mutations (neutral expectation), as well as the total mutation rate (mutation rate per site multiplied by number of sites) on which each SFS has been sampled.
    The total mutation rate for each SFS is thus obtained using the following three-steps procedure.
    First, for each gene, the reference DNA sequence is adjusted with ancestral and fixed polymorphism for each population.
    Second, from this reconstructed sequence, all possible mutations are computed, weighted by the mutation rate between nucleotide estimated at the phylogenetic scale $\UniDimArray{\mu}$.
    Third, these possible mutation are then classified whether they are synonymous or non-synonymous (stop mutations are excluded), non-synonymous mutations are then also classified depending on their selection coefficient ($\Sphy$).
    From the SFS and the total mutation rate for the both synonymous and for the class of non-syonymous, polyDFE estimates the DFE $\phi$ using maximum likelihood.\@.

### 2022-04-13 manuscript/supp-mat.tex
Supp. Mat.

\begin{document}
    \maketitle
    \tableofcontents
    \include{include-figures}
\end{document}

>>>>

\begin{document}
    \maketitle
    \tableofcontents
    \clearpage

    \section{Site-specific mutation-selection codon models}

    \subsection{Summary table}
    \begin{center}
        \begin{longtable}{|l|l|r|r|r|r|}
            \toprule
            Population                     & Species             & Tajima $\theta_{\pi}$ & $\mathbb{P}_{obs}[1<S]$ & $\mathbb{P} [ 1<\beta ]$ & $\mathbb{P} [ 1<\beta  | 1<S]$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{6}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

            \bottomrule
            \endlastfoot
            Diverse (Equus)                & Equus caballus      & $ 0.003$              & $ 0.161$                & $ 0.073$                 & $ 0.779$                     \\
            Diverse (Canis)                & Canis familiaris    & $ 0.004$              & $ 0.069$                & $ 0.041$                 & $ 0.262$                     \\
            Iran (IRBT)                    & Bos taurus          & $ 0.009$              & $ 0.123$                & $ 0.054$                 & $ 0.300$                     \\
            Uganda (UGBT)                  & Bos taurus          & $ 0.010$              & $ 0.139$                & $ 0.060$                 & $ 0.779$                     \\
            Australia (AUCH)               & Capra hircus        & $ 0.004$              & $ 0.105$                & $ 0.061$                 & $ 0.119$                     \\
            France (FRCH)                  & Capra hircus        & $ 0.004$              & $ 0.105$                & $ 0.060$                 & $ 0.059$                     \\
            Iran (IRCA)                    & Capra aegagrus      & $ 0.004$              & $ 0.105$                & $ 0.062$                 & $ 0.053$                     \\
            Iran (IRCH)                    & Capra hircus        & $ 0.005$              & $ 0.093$                & $ 0.058$                 & $ 0.094$                     \\
            Italy (ITCH)                   & Capra hircus        & $ 0.004$              & $ 0.103$                & $ 0.059$                 & $ 0.053$                     \\
            Morocco (MOCH)                 & Capra hircus        & $ 0.005$              & $ 0.077$                & $ 0.057$                 & $ 0.045$                     \\
            Iran (IROA)                    & Ovis aries          & $ 0.009$              & $ 0.107$                & $ 0.048$                 & $ 0.144$                     \\
            Iran (IROO)                    & Ovis orientalis     & $ 0.010$              & $ 0.110$                & $ 0.050$                 & $ 0.177$                     \\
            Iran (IROV)                    & Ovis vignei         & $ 0.009$              & $ 0.127$                & $ 0.049$                 & $ 0.252$                     \\
            Various (ISGC)                 & Ovis aries          & $ 0.009$              & $ 0.108$                & $ 0.049$                 & $ 0.194$                     \\
            Morocco (MOOA)                 & Ovis aries          & $ 0.009$              & $ 0.086$                & $ 0.045$                 & $ 0.144$                     \\
            Barbados                       & Chlorocebus sabaeus & $ 0.004$              & $ 0.104$                & $ 0.072$                 & $ 0.484$                     \\
            Central African Republic (CAR) & Chlorocebus sabaeus & $ 0.007$              & $ 0.105$                & $ 0.066$                 & $ 0.312$                     \\
            Ethiopia                       & Chlorocebus sabaeus & $ 0.006$              & $ 0.097$                & $ 0.068$                 & $ 0.245$                     \\
            Gambia                         & Chlorocebus sabaeus & $ 0.006$              & $ 0.101$                & $ 0.064$                 & $ 0.411$                     \\
            Kenya                          & Chlorocebus sabaeus & $ 0.006$              & $ 0.104$                & $ 0.066$                 & $ 0.323$                     \\
            Nevis                          & Chlorocebus sabaeus & $ 0.004$              & $ 0.105$                & $ 0.068$                 & $ 0.445$                     \\
            South Africa (SA)              & Chlorocebus sabaeus & $ 0.007$              & $ 0.095$                & $ 0.065$                 & $ 0.233$                     \\
            Saint Kitts (SK)               & Chlorocebus sabaeus & $ 0.005$              & $ 0.101$                & $ 0.071$                 & $ 0.442$                     \\
            Zambia                         & Chlorocebus sabaeus & $ 0.007$              & $ 0.101$                & $ 0.070$                 & $ 0.290$                     \\
            African (AFR)                  & Homo sapiens        & $ 0.003$              & $ 0.052$                & $ 0.092$                 & $ 0.779$                     \\
            Ad Mixed American (AMR)        & Homo sapiens        & $ 0.002$              & $ 0.055$                & $ 0.087$                 & $ 0.779$                     \\
            East Asian (EAS)               & Homo sapiens        & $ 0.002$              & $ 0.051$                & $ 0.084$                 & $ 0.779$                     \\
            European (EUR)                 & Homo sapiens        & $ 0.002$              & $ 0.053$                & $ 0.087$                 & $ 0.779$                     \\
            South Asian (SAS)              & Homo sapiens        & $ 0.002$              & $ 0.051$                & $ 0.084$                 & $ 0.779$                     \\
        \end{longtable}
    \end{center}

    \subsection{Correlation with diversity}

    \subsubsection{Proportion of positive selection}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.all_P-Ssup0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.pos_P-Ssup0.scatter.pdf}
    \end{minipage}

    \subsubsection{Proportion of negative selection}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.all_P-Sinf0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.neg_P-Sinf0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.neg-strong_P-Sinf0.scatter.pdf}
    \end{minipage}

    \subsubsection{Proportion of weak selection}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.all_P-Seq0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.neg-weak_P-Seq0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5cat-polyDFE-mC-non-adaptive/regression/results.watterson.pos-weak_P-Seq0.scatter.pdf}
    \end{minipage}

    \input{include-MutSel-non-adaptive}

    \section{SIFT scores}
    \input{include-SIFT-non-adaptive}

    \section{Site-specific codon models.}
    \input{include-Omega-adaptive}
\end{document}
